Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-07-24 GWLDCI.tex
through section 4.1

In \cite{wilsonloop}, the authors show that all admissible Wilson loop diagrams correspond to positroids. That is, they correspond to matroids represented by elements of the positive Grassmannians $\Gr(|\cP|, n)$. Any positroid of rank $k$ on $n$ elements defines a subspace of the positive Grassmannians $\Gr(k, n)$, namely the points which represent it \cite{Postnikov}. These subspaces give a CW structure on $\Gr(k,n)$, with each positroid defining a cell.  \sanote{Susama end here}

\begin{dfn}
Given a Wilson loop diagram $W = (\cP, [n])$, its associated matroid $M(W)$ corresponds to positroid cells in the CW complex on $\Gr(|\cP|, n)$.
\end{dfn}\note{How is this a def as it doesn't define the cells?  Also, should it be cell singular}

With this definition in mind, we have the following corollary:

@@

Now we are ready to prove the main results of this section, namely that two Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}).
We also
give a formula for the number of Wilson loop diagrams in a given equivalence class (Corollary \ref{number of equiv diagrams}), completing the characterization of the correspondence between Wilson loop diagrams and matroids starting in \cite{wilsonloop}.

\begin{thm}\label{same matroid iff equiv}
Let $W= (\cP, [n])$ and $W'= (\cP', [n])$ be two Wilson loop diagrams. They define the same matroid if and only if $W \sim W'$.
\end{thm}

@@

For the if direction, it suffices to prove the result for two equivalent diagrams that differ in exactly one exact subdiagram. Let $W$ and $W'$ be equivalent Wilson loop diagrams which can be decomposed as
\[W = (P \sqcup R, [n]), \qquad W' = (P \sqcup R', [n]),\]
where $(R,V(R))$ and $(R',V(R'))$ are maximal exact subdiagrams of $W$ and $W'$ respectively, and $V(R) = V(R')$.

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the elements of each fiber of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivalence classes.

We know from examples (e.g. \cite{casestudy}) that \hlfix{even after taking the dimension of the positroid cell into consideration}{Sian: I'm not quite sure what ``taking the dimension into consideration'' means here: do we mean ``even after restricting our attention to cells of dimension $3k$''?}\footnote{it is proven in \cite{generalcombinatoricsII} and \cite{Marcott} that Wilson loop diagrams with $k$ propagators map to $3k$ dimensional positroid cells}, many positroid cells do not appear as admissible Wilson loop diagrams with a fixed number of propagators and vertices. That is, the map from admissible Wilson loop diagrams to positroid cells of the correct dimension in the correct Grassmannian neither injective nor surjective. We do not obtain a formula for the number of cells, but do obtain an alternate characterization of the equivalence classes which is the subject of the next section.

@@

Two faces are \emph{parallel} if one can be mapped to the other by a translation in $\mathbb{R}^d$.  In particular parallel faces must be of the same dimension and any two vertices are parallel to each other.

>>>>

In \cite{wilsonloop}, the authors show that each weakly admissible Wilson loop diagrams correspond to a postiroid. That is, each diagram correspond to matroids that can be represented by elements of the positive Grassmannians $\Gr(|\cP|, n)$. Any positroid of rank $k$ on $n$ elements defines a subspace of the positive Grassmannians $\Gr(k, n)$, namely the points which represent it \cite{Postnikov}. These subspaces give a CW structure on $\Gr(k,n)$, with each positroid defining a cell.  In other words, there is a map from weakly admissible Wilson loop diagrams, $W$, to matroids, $M(W)$, to positroid cells in the CW complex on $\Gr(|\cP|, n)$.
\note{How is this a def as it doesn't define the cells?  Also, should it be cell singular} \sanote{got rid of dfn}

With this mapping in mind, we have the following corollary:

@@

Now we are ready to prove the main results of this section, namely that two Wilson loop diagrams define the same matroid (positroid) if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). We also give a formula for the number of Wilson loop diagrams in a given equivalence class (Corollary \ref{number of equiv diagrams}), completing the characterization of the correspondence between Wilson loop diagrams and positroids starting in \cite{wilsonloop}.

\begin{thm}\label{same matroid iff equiv}
Let $W= (\cP, [n])$ and $W'= (\cP', [n])$ be two Wilson loop diagrams. They define the same matroid (positroid) if and only if $W \sim W'$.
\end{thm}

@@

For the if direction, it suffices to prove the result for two equivalent diagrams that differ in exactly one exact subdiagram. Let $W$ and $W'$ be equivalent Wilson loop diagrams which can be decomposed as
\[W = (P \sqcup R, [n]), \qquad W' = (P \sqcup R', [n]),\]
where $(R,V(R))$ and $(R',V(R'))$ are different maximal exact subdiagrams of $W$ and $W'$ respectively, and $V(R) = V(R')$.

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same positroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). If we write \bas M: \textrm{admissible \; WLD} & \rightarrow \textrm{Positroids} \\ W & \rightarrow M(W) \;,\eas the image of this map is given by the distinct equivalence classes. Corollary \ref{number of equiv diagrams} enumerates the elements of each fiber of $M$.  The obvious next question would be to enumerate the number of equivalence classes.

We know from examples (e.g. \cite{casestudy}) that \hlfix{even after taking the dimension of the positroid cell into consideration}{Sian: I'm not quite sure what ``taking the dimension into consideration'' means here: do we mean ``even after restricting our attention to cells of dimension $3k$''?}\sanote{yes, but it may be worth just getting rid of this text.} \footnote{it is proven in \cite{generalcombinatoricsII} and \cite{Marcott} that Wilson loop diagrams with $k$ propagators map to $3k$ dimensional positroid cells}, many positroid cells do not appear as admissible Wilson loop diagrams with a fixed number of propagators and vertices. That is, the map from admissible Wilson loop diagrams to positroid cells of the correct dimension in the correct Grassmannian neither injective nor surjective. We do not obtain a formula for the number of cells, but do obtain an alternate characterization of the equivalence classes which is the subject of the next section.

@@

Two faces are \emph{parallel} if one can be mapped to the other by a translation in $\mathbb{R}^d$.  In particular parallel faces must be of the same dimension. \hlfix{Any two vertices are trivially parallel to each other.}{changed. Correct?}
